---
title: "p8105_hw3_jz3036"
output: github_document
author: "Junyuan Zheng (jz3036)"
date: 2018-10-10
---

* Import necessary packages.
```{r import_packages}
library(tidyverse)
#library(readxl)
```

# Problem 1

* Import the raw data for Problem 1.
```{r data_import_p1}
library(p8105.datasets)
data(brfss_smart2010)
```

* Data manipulation:
```{r data_manipulation_p1}
data_p1 = 
  janitor::clean_names(brfss_smart2010) %>%
  filter(., topic == 'Overall Health') %>%
  filter(., response == 'Poor' | response == 'Fair' | response == 'Good' | response == 'Very good' | response == 'Excellent') %>%
  mutate(., response = factor(response, levels = str_c(c("Excellent", "Very good", "Good", "Fair", "Poor"))))
```

* Q1 In 2002, which states were observed at 7 locations?
```{r p1_q1}
data_p1_q1 = 
  filter(data_p1, year == '2002') %>% 
  group_by(., locationabbr) %>% 
  summarize(., n_loc = n_distinct(locationdesc)) %>% 
  filter(., n_loc == 7)
data_p1_q1
```
CT, FL, and NC were observed at 7 locations.

* Q2 Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
```{r p1_q2}
data_p1 %>% 
  group_by(., year, locationabbr) %>% 
  summarize(., n_loc = n_distinct(locationdesc)) %>% 
  ggplot(aes(x = year, y = n_loc, color = locationabbr)) +
    geom_point() + geom_line() + 
    theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5)) +
    labs(y = 'number of loc', color = 'states', title = 'spaghetti plot')
```

* Q3 Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r p1_q3}
data_p1 %>%
  filter(., year=='2002' | year=='2006' | year=='2010', response=='Excellent') %>% 
  group_by(., year) %>% 
  summarize(., mean = mean(data_value, na.rm = TRUE), SD = sd(data_value, na.rm = TRUE))
```

* Q4 For each year and state, compute the average proportion in each response category (taking the average across locations in a state).
```{r p1_q4_1}
data_p1_q4_1 = 
  group_by(data_p1, year, locationabbr) %>% 
  summarize(., n_loc = n_distinct(locationdesc))

data_p1_q4_2 =
  group_by(data_p1, year, locationabbr, response) %>%
  summarize(., mean_sum = sum(data_value))

data_p1_q4 = left_join(data_p1_q4_2, data_p1_q4_1, by = c('year'='year', 'locationabbr'='locationabbr'))

data_p1_q4 = mutate(data_p1_q4, avg_prop = mean_sum/n_loc)
head(data_p1_q4)
```

* Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r p1_q4_2}
ggplot(data_p1_q4, aes(x = year, y = avg_prop, color = locationabbr)) +
    geom_point() + geom_line() +
    facet_grid(. ~ response) +
    theme(legend.position = "right", plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle=90)) +
    labs(y = 'avg_prop', color = 'states', title = 'state-level averages over time')
```

# Problem 2

* Import the raw data for Problem 2.
```{r data_import_p2}
library(p8105.datasets)
data(instacart)
data_p2 = instacart
```

* write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

* This is a subset of a dataset containing online purchase information for each order. The dataset we are using contains `r dim(data_p2)[1]` rows or purchase of goods, and `r dim(data_p2)[2]` variables for each purchase.
* The variables can be roughly separated into three parts: cumtomer info, goods info, and time info. Variables that could potientially provides useful infomation include the reordered info, order time of the week and the day, and the aisle info.
* For example, the plot below told us people tend to order yogurt on Sunday more, less on the middle of the week.

```{r p2_example}
data_p2 %>%
  filter(., aisle == 'yogurt') %>% 
  group_by(., order_dow) %>% 
  summarize(., n = n()) %>% 
  ggplot(., aes(x = order_dow, y = n)) +
    geom_point() + geom_line() +
    labs(y = '# of orders', x = 'day of a week', title = 'yogurt order trend')
```


* Q1 How many aisles are there, and which aisles are the most items ordered from?
```{r p2_q1}
data_p2_q1 =
  group_by(data_p2, aisle) %>%
  summarize(., n = n()) %>% 
  mutate(aisle_ranking = min_rank(desc(n)))

nrow(data_p2_q1)
filter(data_p2_q1, aisle_ranking == 1)
```
There are `r nrow(data_p2_q1)` different aisles there, in which fresh vegetables are the aisle most ordered from.

* Q2 Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
```{r p2_q2}
data_p2_q1 = arrange(data_p2_q1, desc(n))
ggplot(data=data_p2_q1, aes(x=aisle_ranking, y=n, fill = factor(aisle, levels = aisle))) +
  geom_bar(stat="identity", position = "dodge") +
  scale_x_continuous(breaks = seq(0, 140, by = 20)) +
  scale_y_continuous(breaks = seq(0, 150000, by = 10000)) +
  labs(y = 'number of items ordered', x = 'ranking of items ordered', fill = 'aisles') +
  theme(legend.position = "none", legend.key.size = unit(0.01, 'line'))
```
The legend that is mean to guide finding corresponding aisle is too large, so I didn't print it out here.
Also, the 95th and 96th ranking was tied, so there is a adding up on the plot.

* Q3 Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r p2_q3}
data_p2_q3 =
  filter(data_p2, aisle == 'baking ingredients' | aisle == 'dog food care' | aisle == 'packaged vegetables fruits') %>% 
  group_by(., aisle, product_name) %>% 
  summarize(n = n()) %>% 
  mutate(ranking = min_rank(desc(n))) %>% 
  filter(., ranking == 1) %>% 
  select(., aisle, product_name)
data_p2_q3
```

* Q4 Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 8 table).
```{r p2_q4}
data_p2_q4 = 
  filter(data_p2, product_name == 'Pink Lady Apples' | product_name == 'Coffee Ice Cream') %>% 
  group_by(., product_name, order_dow) %>% 
  summarize(., mean_hour = mean(order_hour_of_day)) %>% 
  spread(., key = order_dow, value = mean_hour)
data_p2_q4  
```

# Problem 3

* Import the raw data for Problem 3.
```{r data_import_p3}
library(p8105.datasets)
data(ny_noaa)
data_p3 = ny_noaa
```

* write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.

* This dataset contains data from weather stations located at NYC region. It contains `r dim(data_p3)[1]` rows or observations, and `r dim(data_p3)[2]` variables for each observation.
* Important variables include Precipatation, Snowfall, Snow depth, max and min temperature.
* A lot of the observation in this dataset have missing value, which could be annoying since a lot of the time when we try to dig info from our data, we need all the variables to be available in each observation. One single missing value could make the one observation useless.

* Q1-1 Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units.

```{r p3_q1-1}
data_p3 = 
  separate(data_p3, date, into = c("year", "month", "day"), sep = "-")
data_p3$tmax = as.integer(data_p3$tmax) 
data_p3$tmin = as.integer(data_p3$tmin)

data_p3 = rename(data_p3, prcp_tenths_of_mm = prcp, snow_mm = snow, snwd_mm = snwd,
       tmax_tenths_of_degrees_C = tmax, tmin_tenths_of_degrees_C = tmin)

head(data_p3)
```

* Q1-2 For snowfall, what are the most commonly observed values? Why?
```{r p3_q1-2}
data_p3 %>%
  group_by(., snow_mm) %>% 
  summarize(., n = n()) %>% 
  mutate(., ranking = min_rank(desc(n))) %>% 
  filter(., ranking == 1)
```
* the most commonly observed snowfall value is 0, because no snowing is more common than snowing.

* Q2 Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r p3_q2}
data_p3 %>%
  filter(., month == '01' | month == '07') %>% 
  group_by(., id, month) %>%
  summarize(., mean_maxt = mean(tmax_tenths_of_degrees_C, na.rm = TRUE)) %>% 
  ggplot(., aes(x = id, y = mean_maxt)) + 
  geom_point() +
    facet_grid(. ~ month) +
    labs(y = 'average max temperature', x = 'observation station') +
    theme(axis.text.x=element_blank())
```
* From the plot above, the average max temperature in Jan is around 0 Celsius, and is around 28 celsius in July. In Jan, there are about three observations of mean of max temperature that are lower than the others.

* Q3 Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r p3_q3}
data_p3$year = as.integer(data_p3$year)
data_p3_1 = 
  group_by(data_p3, year) %>%
  summarize(., mean_maxt = mean(tmax_tenths_of_degrees_C, na.rm = TRUE),
            mean_mint = mean(tmin_tenths_of_degrees_C, na.rm = TRUE)) %>% 
  gather(., key = max_min, value = mean_temperature, mean_maxt, mean_mint) %>%
  ggplot(., aes(x = year, y = mean_temperature, color = max_min)) +
  geom_point() + geom_line() + geom_smooth(method='lm', se = FALSE) +
  scale_x_continuous(breaks = seq(0, 2010, by = 5)) +
  labs(title = 'Mean of Max and Min Temperature across years', y = 'mean temp (tenths of degrees C)')

data_p3_2 =
  filter(data_p3, 0 < snow_mm, 100 > snow_mm) %>% 
  group_by(., year) %>%
  ggplot(., aes(x = year, y = snow_mm, group = year)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(0, 2010, by = 5)) +
  labs(title = 'distribution of snowfall value across years')

library(patchwork)
data_p3_1 / data_p3_2
```